\textsc{\wobble: a data-driven method for precision radial velocities}

### 2019-07-01 paper/paper.tex
submission version
changing model specification

We take the data to be the $M \times N$ matrix Y, where each entry $y_{m,n}$ is the observed logarithmic flux $\ln f$ for pixel $m$ of $M$ at epoch $n$ of $N$. 
We also have a corresponding $M \times N$ matrix of wavelength solutions. 
Again, the entries of this matrix are in logarithmic form ($\ln \lambda$). 
This choice makes the Doppler shift an additive term, considerably simplifying the math to follow.

We model each data column $y_n$ as the sum of the stellar and telluric spectra:
\begin{equation}
y_n = y_{n}^\star + y_{n}^t + noise.
\end{equation}

The stellar spectrum contribution is: 
\begin{equation}
y_{n}^\star = P(v^{obs}_{n}) \mu^{\star},
\label{eqn:star}
\end{equation}
where $\mu^{\star}$ is a spectral template with length $M'$ corresponding to
the length of an arbitrarily-chosen wavelength grid of control points, and
$P$ is a linear operator which applies a Doppler shift by velocity $v_n$ and
interpolates $\mu^\star$ into the observatory rest frame.
Because we work in log-wavelength space, the Doppler shift due to velocity $v$ is simply an additive term:
\begin{equation}
 \ln \lambda(v) = \ln \lambda_{0} + \frac{1}{2} \ln \left(\frac{1 - v/c}{1 + v/c}\right).
\end{equation}
The $P$ operator adds this term to the $M'$-length wavelength grid for the template spectrum before linearly interpolating the template to the $M$-length data space.
$P$ could be represented as a rectangular $M \times M'$ sparse matrix, but the dot product can be implemented efficiently using linear interpolation instead of instantiating the full matrix.

The apparent stellar \RV, $v^{obs}_{n}$, is a sum of the star's actual velocity about its center of mass ($v^{\star}_{n}$) and the projected motion of the Earth about the Solar System barycenter (Barycentric Earth Radial Velocity or \BERV), the latter of which is known.

@@

It is worthy of note that the stellar spectral model shown in Figure \ref{51peg_multispectrum} is an extremely high-\SNR spectrum of 51 Peg. 
This template spectrum is essentially a time-averaged composite spectrum. 
Both it and its residuals as a function of time are scientifically valuable outputs of the \wobble method.

@@

\acknowledgements{It is a pleasure to acknowledge Jacob L. Bean, John M. Brewer, Heather Knutson, Timothy Morton, Melissa Ness, Andreas Seifahrt, and Julian St\"{u}rmer for helpful discussions. 
We are grateful to Hans-Walter Rix and the Max-Planck-Institut f\"ur Astronomie for their hospitality while developing key parts of this project. Special thanks go to Geert Barentsen, Michael Gully-Santiago, Christina Hedges, Susan Mullaly, and Z\'{e} Vin\'{i}cius for their vital contribution to the most important part of \wobble's documentation (the pronunciation guide).

>>>>

We take the data to be the $M \times N$ matrix Y, where each entry $y_{m,n}$ is the observed logarithmic flux $\ln f$ for pixel $m$ of $M$ at epoch $n$ of $N$. 
We also have a corresponding $M \times N$ matrix of wavelength solutions. 
Again, the entries of this matrix are in logarithmic form ($\ln \lambda_{m,n}$). 
\deleted{This choice makes the Doppler shift an additive term, considerably simplifying the math to follow.}

We model each data column $y_n$ as the sum of the stellar and telluric spectra:
\begin{equation}
y_n = y_{n}^\star + y_{n}^t + \textrm{noise}.
\end{equation}

The stellar spectrum contribution is: 
\begin{equation}
y_{n}^\star = P(v^{obs}_{n}) \mu^{\star},
\label{eqn:star}
\end{equation}
where $\mu^{\star}$ is a spectral template with length $M'$ corresponding to
the length of an arbitrarily-chosen wavelength grid of control points, and
$P$ is a linear operator which applies a Doppler shift by velocity $v_n$ and
interpolates $\mu^\star$ into the observatory rest frame.
Because we work in log-wavelength space, the Doppler shift due to velocity $v$ is simply an additive term:
\begin{equation}
 \ln \lambda(v) = \ln \lambda_{0} + D(v) \equiv \ln \lambda_{0} - \frac{1}{2} \ln \left(\frac{1 - v/c}{1 + v/c}\right).
\end{equation}
The $P$ operator adds this term to the $M'$-length wavelength grid for the template spectrum before linearly interpolating the template to the $M$-length data space, \added{so that each predicted value $y_{n,m}^\star$ at data wavelength $\ln \lambda_{m,n}$ is determined by:} \todo{(this is not complete)}
\begin{equation}
y_{m,n}^\star(\ln \lambda_{m,n}) = \mu^{\star}(\ln \lambda_{m'} + D(v^{obs}_{n})).
\label{eqn:Pop}
\end{equation}
$P$ could be represented as a rectangular $M \times M'$ sparse matrix, but the dot product can be implemented efficiently using linear interpolation instead of instantiating the full matrix.

The apparent stellar \RV, $v^{obs}_{n}$, is a \replaced{sum}{combination} of the star's actual velocity \replaced{about its center of mass}{in the barycentric reference frame} ($v^{\star}_{n}$) and the projected motion of the Earth about the Solar System barycenter (Barycentric Earth Radial Velocity or \BERV), the latter of which is known. 
\todo{(add Wright equation!)}

@@

It is worthy of note that the stellar spectral model shown in Figure \ref{fig:51peg_spectrum} is an extremely high-\SNR spectrum of 51 Peg. 
This template spectrum is essentially a time-averaged composite spectrum. 
Both it and its residuals as a function of time are scientifically valuable outputs of the \wobble method.

@@

\acknowledgements{\added{We gratefully acknowledge the referee, Ian Czekala, for a thoughtful, constructive, and helpful review.}
It is a pleasure to acknowledge Jacob L. Bean, John M. Brewer, Heather Knutson, Timothy Morton, Melissa Ness, Andreas Seifahrt, and Julian St\"{u}rmer for helpful discussions. 
We are grateful to Hans-Walter Rix and the Max-Planck-Institut f\"ur Astronomie for their hospitality while developing key parts of this project. Special thanks go to Geert Barentsen, Michael Gully-Santiago, Christina Hedges, Susan Mullaly, and Z\'{e} Vin\'{i}cius for their vital contribution to the most important part of \wobble's documentation (the pronunciation guide).
